---
title: "L02 ggplot 1"
subtitle: "Data Visualization (STAT 302)" 
author: "Shuo Han"
format:
  html:
    toc: true
    self-contained: true
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
---



## Goals

1. use additional color, shape, size, transparency aesthetics
2. recognize a variety of different geom layers
3. set x and y limits
4. read documentation and access help files
5. read in .rda and .txt files
6. use proper organization 

::: callout-tip
## Organization is important!

1.  All packages should always be loaded at the TOP of the document.

2.  All data should be in your `/data` subdirectory. Data loading should go at the top of the document under the libraries.
:::

## Datasets

For this activity we will be using the `blue_jays`, `BA_degrees`, `dow_jones_industrial` and `cdc` data set that should be located in your `/data` subdirectory. An explanation of the variables in the data set are provided below in their respective Exercises.


```{r}
#| label: load-packages
#| message: false
# Load package(s)
library(splines)
library(tidyverse)
library(dplyr)
library(forcats)
library(lubridate)
# Load data
load("data/blue_jays.rda")
load("data/BA_degrees.rda")
load("data/dow_jones_industrial.rda")
cdc = read.table("data/cdc.txt",sep = "|", header=TRUE)
```


## Exercise 1

We'll be using data from the `blue_jays.rda` dataset to answer the following questions and construct the following plots. 

Below is a description of the variables contained in the dataset.

* `BirdID` - ID tag for bird
* `KnownSex` - Sex coded as F or M
* `BillDepth` - Thickness of the bill measured at the nostril (in mm)
* `BillWidth` - Width of the bill (in mm)
* `BillLength` - Length of the bill (in mm)
* `Head` - Distance from tip of bill to back of head (in mm)
* `Mass` - Body mass (in grams)
* `Skull` - Distance from base of bill to back of skull (in mm)
* `Sex` - Sex coded as `0 = female` or `1 = male`

<br>

### Question 1.1

Three commonly used aesthetics for enhancing a graphic are color, size, and shape. We can either map variables or fixed values to these.

In the `blue_jays` dataset which variables would be appropriate to map to the `size` aesthetic? The `shape` aesthetic? The `color` aesthetic?

**Answer--** For variables KnownSex and Sex, we can use the'shape'to tell the proportion.
For variables BillDepth, BillWidt and BillLength, we can use the size. It's easier to indicate the numbers.
<br>

### Plot 1

Construct a scatterplots of `Head` by `Mass` with the following criteria:

- `color` aesthetic set to Northwestern purple (`#4E2A84`)
- `shape` aesthetic set a solid/filled triangle
- `size` aesthetic set to 2.
```{r}
#| lable: head-mass
#| message: false
#| warning: false

ggplot(data = blue_jays, mapping = aes(x = Mass, y = Head))+
  geom_point(pch=17,color = "#4E2A84",size = 2)+
  labs(x="Mass",
       y="Head",
       title ="Head by Mass"
       )+ 
   theme_minimal()
```

<br>

### Plot 2

Construct a scatterplot of `Head` by `Mass` with the following criteria:

- `size` aesthetic to 2.
- Either `Sex` or `KnownSex` mapped to the `color` aesthetic.
```{r}
#| lable: head-mass-sex
#| message: false
#| warning: false

ggplot(data = blue_jays, mapping = aes(x = Mass, y = Head, color = KnownSex))+
  geom_point(size = 2)+
  labs(x="Mass",
       y="Head",
       title ="Head by Mass with sex"
       )+ 
   theme_minimal()

ggplot(data = blue_jays, mapping = aes(x = Mass, y = Head, color = Sex))+
  geom_point(size = 2)+
  labs(x="Mass",
       y="Head",
       title ="Head by Mass with sex"
       )+ 
   theme_minimal()
```

<br>

### Question 1.2

Was it more appropriate to map `Sex` or `KnownSex` to the `color` aesthetic in Plot 2 and why?

**Answer--** I think it's more appropriate to use KnownSex, because the F and M in the plot makes the information more clear.
<br>

### Question 1.3

Consider the `color` aesthetic in Plot 1 and Plot 2 above. **Explain why** these two usages of the `color` aesthetic are meaningfully different. 

**Answer--** Plot 1: The color tells nothing and has no connection with the data. But for plot 2: The color tells is connected with the Sex of the bird and we can know the sex of a bird from the color.

<br>

## Exercise 2

We'll be using data from the `BA_degrees.rda` which is already in the `/data` subdirectory in our **data_vis_labs** project. Below is a description of the variables contained in the dataset.

`BA_degrees.rda`

* `field` - field of study
* `year_str` - academic year (e.g. 1970-71)
* `year` - closing year of academic year
* `count` - number of degrees conferred within a field for the year
* `perc` - field's percentage of degrees conferred for the year


```{r}
#| label: ex-2-data-wrangle

# Wrangling for plotting
ba_data <- BA_degrees %>%
  # mean % per field
  group_by(field) %>%
  mutate(mean_perc = mean(perc)) %>%
  # Only fields with mean >= 5%
  filter(mean_perc >= 0.05) %>%
  # Organizing for plotting
  arrange(desc(mean_perc), year) %>%
  ungroup() %>%
  mutate(field = fct_inorder(field))
```

<br>

Using the `ba_data` dataset derived from `BA_degrees`, recreate the following graphics as precisely as possible.

### Plot 1
```{r}
#| lable: Plot 1
#| message: false
#| warning: false

ggplot(data = ba_data, mapping = aes(x = year, y = perc))+
  facet_wrap(~field)+
  geom_line()+
  labs(x="Year",
       y="Proportion of Degree",
       title ="Plot 1"
       )
```

<br>

### Plot 2

*Hints:*

- Transparency is 0.5
- Color used is`"red"`
```{r}
#| lable: Plot 2
#| message: false
#| warning: false

ggplot(data = ba_data, mapping = aes(x = year, y = perc))+
  facet_wrap(~field)+
  geom_line()+
  geom_area(alpha=0.5, fill="red")+
  labs(x="Year",
       y="Proportion of Degree",
       title ="Plot 2"
       )
```

<br>

### Plot 3
```{r}
#| lable: Plot 3
#| message: false
#| warning: false

ggplot(data = ba_data, mapping = aes(x = year, y = perc, color = field))+
  geom_line()+
  labs(x="Year",
       y="Proportion of Degree",
       title ="Plot 3"
       )
```


<br>

## Exercise 3

Using the `dow_jones_industrial` dataset, track the Dow Jones Industrial Average at close each day during 2009. Recreate the following graphics as precisely as possible.


```{r}
#| label: ex-3-data-wrangle

# Only keep observations within 2009 range
djia_date_range <- dow_jones_industrial %>%
  filter(
    date >= ymd("2009/01/01") & date < ymd("2010/01/01")
    )
```

<br>

### Plot 1

It will be easier to see the trend if we add a smooth fit line to the graph.

*Hints:*

- colors: "purple", "red", "green"
- Wiggliness for `loess` is 0.3
```{r}
#| lable: Plot 1
#| message: false
#| warning: false

ggplot(data = djia_date_range, mapping = aes(x = date, y = close))+
  geom_line(color ='purple')+
  geom_smooth(color = "green", span = 0.3, fill="red")+
  labs(x = "",
       y="Dow Jones Industrial Average",
       title ="Plot 1"
       )
```

<br>

### Plot 2

There are many different smoothing methods in statistics. By default `geom_smooth()` uses `loess`. 

Another extremely popular smoothing method in statistics is the use of splines. A spline function is a curve constructed from polynomial segments that are subject to conditions or continuity at their joints. Generally, a difficulty is you need to determine the number of `knots`, however this method is popular in predictive models.

Add a natural cubic spline smoother with `df = 6`.

*Hints:*

- need the `splines` package
- use method `"lm"` 
- `y ~ ns(x, 6)`
```{r}
#| lable: Plot 2
#| message: false
#| warning: false

ggplot(data = djia_date_range, mapping = aes(x = date, y = close))+
  geom_line()+
  geom_smooth(method = lm, formula = y ~ splines::ns(x, 6), se = FALSE)+
  labs(x = "",
       y="Dow Jones Industrial Average",
       title ="Plot 2"
       )
```

<br>

## Exercise 4

We'll be using the BRFSS (Behavioral Risk Factor Surveillance System) survey collected annually by the Centers for Disease Control and Prevention (CDC). The data can be found in the provided `cdc.txt` file --- place this file in your `/data` subdirectory. The dataset contains 20,000 complete observations/records of 9 variables/fields, described below.

* `genhlth` - How would you rate your general health? (excellent, very good, good, fair, poor)
* `exerany` - Have you exercised in the past month? (`1 = yes`, `0 = no`) 
* `hlthplan` - Do you have some form of health coverage? (`1 = yes`, `0 = no`) 
* `smoke100` - Have you smoked at least 100 cigarettes in your life time? (`1 = yes`, `0 = no`) 
* `height` -  height in inches
* `weight` - weight in pounds
* `wtdesire` - weight desired in pounds
* `age` - in years
* `gender` - `m` for males and `f` for females

<br>

Using the `cdc` dataset, we want to visualize the **distribution** between weight and gender based on general health. Recreate the Plot 1 and Plot 2 which represent 2 attempts at this. 

<br>

### Plot 1

*Hints:*

- Transparency is 0.2
- Horizontal axis should have lower limit of 50 and upper limit of 400
```{r}
#| lable: Plot 1
#| message: false
#| warning: false

cdc_new = cdc %>% 
  mutate(
    genhlth = factor(genhlth, levels = c("excellent","very good","good","fair","poor"))
  )
ggplot(cdc_new,aes(weight, color = genhlth,fill = genhlth)) +
  geom_density(alpha = 0.2, na.rm = TRUE) + 
  xlim(50, 400) + 
  facet_wrap(~gender)
```

<br>

### Plot 2

*Hints:*

- Transparency is 0.4
```{r}
ggplot(cdc, aes(gender, weight,fill=gender)) + 
  geom_boxplot(alpha=0.4)+
  facet_grid(~genhlth)
```
<br>

### Plot 3

Choose one more way to visualize the **distribution** of weight based on gender and general health. (Many ways to do this). 

*Requirements:*

- The geom layer cannot be geom_boxplot() or geom_density()
- Appropriately label the x- and y- axis
- Add a title
```{r}
ggplot(cdc, aes(gender,weight,color=gender,fill=gender)) + 
  facet_grid(~genhlth)+
  geom_bar(stat = "summary_bin", fun = mean)+
  labs(
    x="Gender",
    y = "Average Weight",
    title = "Average weight under different gender and genhlth"
  )
```

<br>

### Question 4.1

Compare and contrast the different ways to visualize a distribution. In other words, when might you use a histogram vs density vs boxplot? Which did you find the most useful in interpreting the plots above?

**Answer--**For the boxplot, it is mainly used to reflect the characteristics of the original data distribution, and can also be used to compare the distribution characteristics of multiple groups of data. And histogram is related to density, we can estimate the density through histogram given enough data. I like density better, the information is more readable.

<br>

## Exercise 5

Using the `cdc` dataset, construct this plot in **two** ways. Once using `geom_bar()` and once using the `geom_col()`. May need to do some data wrangling for one of these.

```{r}
ggplot(cdc, aes(genhlth)) + 
  geom_bar()+
  xlim(c("excellent","very good","good","fair","poor"))+
  labs(
    y = "Count",
    x = "General Health" 
  )
```


```{r}
cdc2 <- cdc %>%
  group_by(genhlth) %>%
  mutate(count = n())
cc=cdc2[c("genhlth","count")] %>% 
  distinct()
ggplot(cc, aes(x=genhlth,y=count)) + 
  geom_col()+
  xlim(c("excellent","very good","good","fair","poor"))+
  labs(
    y = "Count",
    x = "General Health" 
  )
```

