---
title: "L03 ggplot II"
subtitle: "Data Visualization (STAT 302)" 
author: "Shuo Han"
format:
  html:
    toc: true
    self-contained: true
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
---



## Overview

The goal of this lab is to continue the process of unlocking the power of `ggplot2` through constructing and experimenting with a few basic plots.

## Datasets

We will be using the BRFSS survey which was introduced in the last lab. The data was supplied in `cdc.txt` file and should be in your `/data` subdirectory. If not, you need to download the last lab to get this data file. As a reminder, the dataset contains 20,000 complete observations/records of 9 variables/fields, described below.

* `genhlth` - How would you rate your general health? (excellent, very good, good, fair, poor)
* `exerany` - Have you exercised in the past month? (`1 = yes`, `0 = no`) 
* `hlthplan` - Do you have some form of health coverage? (`1 = yes`, `0 = no`) 
* `smoke100` - Have you smoked at least 100 cigarettes in your life time? (`1 = yes`, `0 = no`) 
* `height` -  height in inches
* `weight` - weight in pounds
* `wtdesire` - weight desired in pounds
* `age` - in years
* `gender` - `m` for males and `f` for females

```{r}
# Load package(s) 
library(tidyverse)
library(dplyr)
library(forcats)
# Data
cdc = read.table("data/cdc.txt",sep = "|", header=TRUE)
```

## Exercise 1

Using the `cdc` dataset, we want to look at the relationship between height and weight. Recreate the following graphics as precisely as possible.

<br>

### Plot 1

*Hints:*

- Transparency is 0.2
- Minimal theme

```{r}
#| lable: plot1
#| message: false
#| warning: false
ggplot(cdc, aes(height, weight,color = gender)) + 
  geom_point(alpha=0.2) + 
  geom_smooth(method = lm, fullrange= TRUE)+
  labs(
    x = "Height(in)",
    y = "Weight(lbs)"
  )+
  theme_minimal()
```

<br>

### Plot 2

*Hints:*

- size = 0.7
```{r}
ggplot(cdc, aes(height, weight,color = gender)) + 
  geom_density_2d(size = 0.7)+
  labs(
    x = "Height(in)",
    y = "Weight(lbs)"
  )+
  theme_minimal()
```


<br>

### Plot 3

*Hints:*

- `bins` set to 35
```{r}
ggplot(cdc, aes(height, weight)) + 
  geom_hex(bins = 35)+
  labs(
    x = "Height(in)",
    y = "Weight(lbs)"
  )+
  theme_minimal()
```


<br>

### Plot 4

*Hints:*

- use a stat layer, not a geom layer
- `geom = "polygon"` 
```{r}
ggplot(cdc,aes(x = height, y = weight)) +
stat_density2d(aes(fill = ..level..),geom = "polygon",show.legend = FALSE) +
facet_wrap(~ gender) +
labs(
x = "Height (in)",
y = "Weight (lbs)"
) +
theme_minimal()
```


<br>

## Exercise 2

Using the `cdc_means` dataset derived from the `cdc` dataset, recreate the following graphic as precisely as possible.

*Hints:*

- Hex color code `#56B4E9`
- 95% confidence intervals (1.96 or `qnorm(0.975)`)
- Some useful values: 0.1, 0.7
```{r}
# data wrangling
# calc mean and se for CI
cdc_means <- cdc %>%
  mutate(wtloss = weight - wtdesire) %>%
  group_by(genhlth) %>%
  summarize(
    mean = mean(wtloss),
    se = sd(wtloss) / sqrt(n())
  ) %>%
  mutate(genhlth = fct_reorder(factor(genhlth), desc(mean))) 


```

```{r}

ggplot(cdc_means,aes(genhlth, mean))+
  geom_col(fill = "#56B4E9", width = 0.7)+
  geom_errorbar(aes(ymin = mean - se*qnorm(0.975),
                    ymax = mean + se*qnorm(0.975)), width = 0.1)+
 theme_minimal()+
 labs(
   x = "General Health",
   y = "Mean desired weight loss in lbs"
 )
```

<br>


<br>

## Exercise 3

Using the `cdc_weight_95ci` dataset derived from the `cdc` dataset, recreate the following graphic as precisely as possible.

*Hints:*

- Useful values: 0.1, 0.5
- Need to know CI formula

```{r}
#| label: ex-3-plot

# data wrangling
# calculate mean, se, and margin of error for CI formula
cdc_weight_95ci <- cdc %>%
  mutate(
    genhlth = factor(genhlth, levels = c("excellent","very good","good","fair","poor"))
  ) %>% 
  group_by(genhlth, gender) %>%
  summarise(
    mean_wt = mean(weight),
    se = sd(weight) / sqrt(n()),
    moe = qt(0.975, n() - 1) * se
  ) %>% 
  ungroup()
```

```{r}

ggplot(cdc_weight_95ci,aes(mean_wt, gender, color = genhlth))+
  geom_errorbar(aes(xmin = mean_wt - moe,
                    xmax = mean_wt + moe),width = 0.1, position = position_dodge(0.5))+
  geom_point(position = position_dodge(0.5))+
  theme_minimal()+
    labs(
    x= "Weight(lbs)",
    y = "Gender")
```
