---
title: "L08 Coords"
subtitle: "Data Visualization (STAT 302)" 
author: "Shuo Han"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: show
    link-external-newwindow: true
    fig-dpi: 300
execute:
  warning: false
from: markdown+emoji 
editor_options: 
  chunk_output_type: console
---

## Overview

The goal of this lab is the use of an alternative coordinate system in `ggplot2` to build a plot. 

## Datasets

We'll be using the `mod_nba2014_15_advanced.txt` dataset --- add to your project's `/data` subdirectory. The `codebook_mod_nba2014_15_advanced.txt` provides a quick description of the variables in the dataset --- suggest adding it to the `/data` subdirectory as well. 

```{r}
#| label: load-pkgs-data
#| message: false

# load package(s)
library(tidyverse)
# load data
nba_dat = read_delim("data/mod_nba2014_15_advanced.txt", delim = "|") %>% 
  janitor::clean_names()
```

## Exercise 

<!-- See the html instructions since they are complete with images and are the official instructions. -->

Using the `mod_nba2014_15.txt` dataset we will try to recreate/approximate the plot type featured in the [http://fivethirtyeight.com/](fivethirtyeight.com) article [Kawhi Leonard Is The Most Well-Rounded Elite Shooter Since Larry Bird](http://fivethirtyeight.com/features/kawhi-leonard-is-the-most-well-rounded-elite-shooter-since-larry-bird/) for any player of your choice for the 2014-2015 season. 

**Replicate the graphic below as closely as possible:**

### Start with data wrangling

When data wrangling we will need a helper function for creating the quartiles for players. Additionally, it will be useful to exclude players that played less than 10 games or played less than 5 minutes a game. That is, we only want to include "qualified" players.

```{r}
#| label: helper-function

# Returns quartile rank
quartile_rank <- function(x = 0:99) {
  # Set quartile
  quart_breaks <- c(
    -Inf,
    quantile(x,
      probs = c(.25, .5, .75),
      na.rm = TRUE
    ),
    Inf
  )
  cut(x = x, breaks = quart_breaks, labels = FALSE)
}
```

:::{.callout-important}
**Add comments to** the code below where indicated. The added comments should concisely describe what the following line(s) of code do in the data wrangling process  
:::

```{r}
#| label: data-wrangling
# Graphical data
nba_graph_dat <- nba_dat %>% 
  # Remove the player who behaved badly
  filter(g >= 10, mp / g >= 5) %>% 
  # Use function quartile rank to get the quartile rank
  # for 4 variables
  mutate(
    ts_quant = quartile_rank(ts_perc),
    trb_quant = quartile_rank(trb_perc),
    dbpm_quant = quartile_rank(dbpm),
    ast_quant = quartile_rank(ast_perc),
    usg_quant = quartile_rank(usg_perc)
  ) %>% 
  # Only keep the _quant variable for graphing
  select(player, contains("_quant")) %>% 
  # Group the data by players, _quant and quant levels
  pivot_longer(
    cols = -player, 
    names_to = "variable", 
    values_to = "value"
    ) %>% 
  # Sort the date by players names
  arrange(player)
```

### Build the plot

*Hints:*

- Image width is 8 in, height is 10 in --- use code chunck yaml
- center the image output --- use code chunck yaml
- shrink size down for display to 50% width --- use code chunck yaml
- All added text, except title, is done with `annotate()`
- set `axis.text` to blank in theme
- size 3 and 5 could be useful

```{r}
#| label: Base_plot
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 10
#| out-width: "50%"

# id player for chart
player_name = "LeBron James"
# build plot
nba_graph_dat %>%
  filter(player == "LeBron James") %>%
  ggplot(aes(x = variable, y = value)) +
  geom_col(width = 1, fill = "#F07786") +
  # placing dotted lines
  geom_hline(yintercept = 1:4, linetype = "dotted") +
  geom_segment(x = 0.5:4.5, y = 0, xend = 0.5:4.5, yend = 4) +
  scale_x_discrete(
    name = NULL,
    expand = c(0, 0),
    limits = c("ts_quant", "usg_quant", "dbpm_quant", "trb_quant", "ast_quant")
  ) +
  scale_y_continuous(
    name = NULL,
    expand = c(0, 0)
  ) +
  theme(
    axis.text = element_blank()
  ) +
ggtitle(str_c(player_name, "(2015)", sep = "\n"))+
  # sector labels
  annotate(
    geom = "text", 
    label = c(
      "TRUE\nSHOOTING", "USAGE\nRATE", 
      "DEFENSIVE\nBPM", "REBOUND\nRATE",
      "ASSIST\nRATE"
      ),
    x = 1:5,
    y = 5,
    size = 5
  ) +
  annotate(
    geom = "text",
    x = 3, y = 0.7:3.7, 
    label = c("1st-25th", "25th-50th", "50th-75th", "75th-99th"),
    size = 3) +
  coord_polar() +
  theme(
        axis.text = element_blank(), 
        panel.grid = element_blank())
```

## Challenge

<!-- See the html instructions since they are complete with images and are the official instructions. -->

**Challenge is optional for all students, but we recommend trying them out!** 

Continue the work of trying to recreate the graphic from the article. There is a pre built theme from `ggthemes` that might be useful. Your final plot should look as follows:
