---
title: "Introduction to {ggplot2}"
subtitle: "Data Visualization (STAT 302)"
author: "YOUR NAME"

format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true
    fig-dpi: 300
    fig-align: center

execute:
  echo: false
  warning: false

from: markdown+emoji 
---

## Overview

This activity is meant to pull together all the data visualization skills and knowledge we have built using `ggplot2` over the last several weeks. Ideally, students should feel comfortable with being able to build the **Level Up Bar Plot**. While the **Advanced Bar Plot** might not be immediately attainable, students should have the foundation knowledge to be able to figure out how to build it. We cannot teach you everything, but we hope to have prepared you to be able to learn and grow on your own.

### Citation

This activity was adapted from a workshop given by Tanya Shapiro.

## Load pacakges and data 

The data we're using today is from [**The Movie Database**](https://www.themoviedb.org/) and contains information on over 35K **Horror Movies** from the 1950s until today 😱

```{r}
#| label: load-pkgs-data

# load packages
library(tidyverse)
library(ggimage)
library(sysfonts)
library(showtext)

# load fonts
font_add_google("Roboto","Roboto")
font_add_google("Creepster","Creepster")

# load data
raw <- read_csv("data/horror_movies.csv")

# to automatically render fonts, use showtext_auto()
showtext::showtext_auto()
```



### Quick Clean-Up

We'll use **dplyr** to make a few tweaks, namely to modify existing columns and add a couple of new ones with `mutate`. To learn more about dplyr, check out this slick [cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf) by RStuido.

🧹 To-Dos:

-   Express *budget* and *revenue* in millions
-   Convert *release_date* from char to date
-   Create *release_year* and *release_month*
-   Append *poster_path* to themoviedb url to create a *poster_url*

```{r}
#| label: data-cleanup

# url for poster images, need this to concatenate with poster path
base_url <- 'https://www.themoviedb.org/t/p/w1280'

# build basic dataset
df <- raw %>% 
  mutate(
    budget = budget/1000000,
    revenue = revenue/1000000,
    #convert release date from char to date
    release_date = as.Date(release_date),
    #truncate date on year and month
    release_year = as.numeric(format(release_date, '%Y')),
    release_month = as.numeric(format(release_date, '%m')),
    #concatenate base url with poster path to get poster url
    poster_url = paste0(base_url,poster_path)
   )
```

## Bar Plot

Visualize the top 10 horror movies by revenue.

### Pre-Plotting

To rank our data set in descending order by revenue, we can use dplyr `arrange()`. We'll simplify and pick the columns we need for plotting with `select()`. 

```{r}
#| label: bar-plot-data

# get top movies based on revenue
df_top_movies<- df %>% 
  #arrange data in descending order revenue
  arrange(-revenue) %>% 
  #select specific columns
  select(title, release_year, revenue, budget, poster_url) %>% 
  #select first 10 rows
  slice_head(n = 10)
```

### Basic Bar Plot

```{r}
#| label: basic-bar-plot
#| fig-width: 8
#| fig-height: 8

# plot horizontal bar chart, add plot labels
ggplot(data = df_top_movies, mapping = aes(y = title, x = revenue)) +
  geom_col() +
  labs(
    x = "Revenue (millions)",
    y = NULL,
    title = "Horror Movies That Killed The Box Office",
    subtitle = "Top horror movies based on total revenue."
  )
```

### Level Up Bar Plot

```{r}
#| label: interm-bar-plot
#| fig-width: 8
#| fig-height: 8

# set some palette colors
pal_bar <- "#A70000" # fill for bars
pal_bg <- '#191919' # background
pal_text <- "white" # text color
pal_subtext <- "#DFDFDF" # text color
pal_grid <- "grey30" # grid color

# build plot
ggplot(
  data = df_top_movies,
  mapping = aes(y = reorder(title, revenue), x = revenue)
)+
  geom_col(fill = pal_bar)+
  labs(
    x = "Revenue (millions)",
    y = NULL,
    title = "Horror Movies That killed the Box Office",
    subtitle = "Top horror movies based on total revenue"
  )+
  theme(
    panel.background = element_rect(fill = pal_bg),
    plot.background = element_rect(fill = pal_bg),
    axis.text = element_text(color = pal_text),
    axis.ticks = element_blank(),
    plot.title = element_text(color = pal_text),
    plot.subtitle = element_text(color = pal_text),
    panel.grid = element_blank(),
    panel.grid.major.x = element_line(color = pal_grid)
  )
```

### Advanced Bar Plot

```{r}
#| label: adv-bar-plot
#| fig-width: 8
#| fig-height: 8

# build plot
```
